---
title: "MidtermProjectPart2"
output: html_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("enrichR")
#install.packages("openxlsx")# independant of java
#install.packages("devtools") 
#devtools::install_github("wjawaid/enrichR")
library(enrichR)
library(openxlsx)
```

## Import T-test results so that we can short list further
```{r}
#folder that contains group comparison results
fileName <- "output/TTest_Severe_(Comp).vs._Control_(Base).TTest.csv"

# Read in group comparison results
result1 <- read.csv(file = fileName, 
                        header = T, stringsAsFactors = F, row.names = 1)
```

#Filtering/shortlisting 
Short list results based in p-value cut off (pvalue <= 0.01)
```{r}
pValueCutOff <- 0.01
which1 <- ((as.numeric(result1$Pvalue) <= pValueCutOff))
table(which1) #146 genes with p <= 0.01

resultShort <- result1[which1, ] #short listed genes 
```

## Export the short listed results for reference
```{r}
#length of short listed results
nrow(resultShort)

write.csv(resultShort, file = "output/TTest_results_shortlist2.csv")
```

## Clean gene names
Note - the gene names are in the form of "reporterid_x_at|genename". So need to split this.
```{r}
funcSplit <- function(rep_gene) {
    rep_gene_split <- unlist(strsplit(x = rep_gene, 
                                      split = "|", 
                                      fixed = TRUE))
    gene <- rep_gene_split[2]
    return(gene)
}
geneListSplit <- apply(X = as.matrix(row.names(resultShort)), 
                       MARGIN = 1, FUN = funcSplit )
head(geneListSplit) #cleaned gene names

#remove duplicates
geneListSplit1 <- unique(geneListSplit)

# remove NA value
geneListFinal <- na.omit(geneListSplit1) #  genes

#print number of unique genes
length(geneListFinal)
write.table(x = geneListFinal, 
          file = "output/shortListedUniqueGenes.tsv",
          quote = F, sep = "\t", row.names = FALSE, col.names = F)


```

## Load Databases for Enrichr R package , and check connection
```{r}
#checking if EnrichR website and packing are working
dbs <- enrichR::listEnrichrDbs()  #total number of databases available = 200+ 

#testing if EnrichR package is working
testOutput <- enrichR::enrichr(genes = c("Runx1", "Gfi1", "Gfi1b", "Spi1", "Gata1", "Kdr"), databases = "KEGG_2021_Human")
head(testOutput[[1]])

#List of databases for which enrichment analysis will be run
dblist1 <- read.csv(file = "input/2023-EnrichR-Databases.txt", header = F, stringsAsFactors = F)

head(dblist1)
```

## Call function to run Enrichment

```{r}
# set output file name
outputFileName1 <- paste("output/YukthaPenumala_EnrichR_MidtermProject.xlsx", sep="")

#Load R script into the environment
source(file = "functionEnrichment.R") 

#call function to run Enrichment
functionEnrichment(dblist1, geneListFinal, outputFileName1)
```
